---
title: "Viz_and_Eda"
author: "Dionna Attinson"
date: "9/26/2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
library(patchwork)
``` 

## create the weather data
```{r load_data, cache = TRUE}
weather_df = 
  rnoaa:: meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

## create a ggplot

```{r}
ggplot(weather_df, aes(x= tmin, y= tmax)) +
  geom_point()
```

alternate way of making this plot 

```{r}
weather_df %>% 
  ggplot(aes(x= tmin, y = tmax)) +
  geom_point()
```

saving initial plots 

```{r} 
weather_df %>% filter(name == "CentralPark_NY")
```

```{r}
scatterplot= 
  weather_df %>% 
  ggplot(aes(x= tmin, y= tmax)) +
  geom_point()

scatterplot
```

adding color...

```{r}
weather_df %>% 
    ggplot(aes(x= tmin, y= tmax)) +
    geom_point(aes(color=name), alpha = .4)
```

vs

```{r}
weather_df %>% 
    ggplot(aes(x = tmin, y= tmax, color = name)) + 
    geom_point(aes(color=name), alpha = .5) + 
    geom_smooth(se=FALSE)
```

facet!
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

this is fine but not interesting 

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size=prcp), alpha = .5) +
  geom_smooth(size = 2, se = FALSE)
```

## Some extra stuff

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_smooth(size = 2, se = FALSE)
```

2d density

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_bin2d() 
```

## More kinds of plots!! 

Understand the distribution of a single variable 

```{r}
weather_df %>% 
    ggplot(aes(x = tmax, fill = name)) +
    geom_histogram()
```


```{r}
weather_df %>% 
    ggplot(aes(x = tmax, fill = name)) +
    geom_histogram(position = "dodge")
```

```{r}
weather_df %>% 
    ggplot(aes(x = tmax, fill = name)) +
    geom_histogram(position = "dodge") +
    facet_grid(~name)
```

## A Density Plot 
```{r}
weather_df %>% 
    ggplot(aes(x = tmax, fill = name)) +
    geom_density(alpha = .3) 
```

## A Box Plot
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_violin()
```

Ridge plots!!

```{r}
weather_df %>% 
  ggplot(aes(x= tmax, y= name)) + 
  geom_density_ridges()
```


##saving a plot 

```{r} 
ggp_ridge_temp = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = name)) + 
  geom_density_ridges()

ggsave("ggplot_temp_ridge.pdf", ggp_ridge_temp)
```

```{r, fig.width = 6}
weather_df %>% 
    ggplot(aes(x = tmin, y = tmax, color = name)) +
    geom_point(alpha = .4) + 
    geom_smooth(se = FALSE)
```

```{r, fig.width = 12}
weather_df %>% 
    ggplot(aes(x = tmin, y = tmax, color = name)) +
    geom_point(alpha = .4) + 
    geom_smooth(se = FALSE)
```


```{r}
weather_df %>% 
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_point(alpha = .5) +
  labs(
      title = "temperature plot",
      x = "minimum temp (c)",
      y = "maximum temp (c)", 
      caption = "Data from NOAA via rnoaa package"
  ) + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ยบ C", "0", "15")
  ) +
  scale_y_continuous(
    trans = "sqrt", 
    position = "right")
```

## Color themes produced
```{r}
ggp_base = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE)
ggp_base
```

## Other themes 
```{r}
ggp_base + 
  theme_minimal() + 
  theme(legend.position="bottom")
```

Where do my warnings come from...

```{r}
weather_df %>% 
  filter(is.na(tmin))
```

## more than one dataset

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name== "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
```

## Multi-panel Plots
```{r}
# install.packages("devtools")
devtools::install_github("thomasp85/patchwork")
```
```{r}
ggp_scatter = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

ggp_density = 
  weather_df %>% 
  ggplot(aes(x = tmin))+ 
  geom_density()

ggp_box = 
  weather_df %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_boxplot()

(ggp_scatter + ggp_density) /ggp_box
```

## Data Manipulation
reorder function organizes data by size
```{r}
  weather_df %>% 
    mutate(
      name = factor(name),
      name = fct_reorder(name, tmax)
    ) %>% 
  ggplot(aes(x = name, y = tmax, color = name)) + 
  geom_boxplot()
```

```{r} 
pup_data = 
  read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  mutate(wt_gain = gd18_weight - gd0_weight,
         day_of_tx = as.numeric(day_of_tx))

fas_data = left_join(pup_data, litter_data, by = "litter_number") 

fas_data %>% 
  select(sex, dose, day_of_tx, pd_ears:pd_walk) %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day") %>% 
    drop_na(dose, day_of_tx, pn_day) %>% 
  mutate(
    outcome = factor (outcome),
    outcome = fct_reorder(outcome, pn_day)
  ) %>%
  mutate(outcome = forcats::fct_reorder(outcome, day_of_tx, median)) %>% 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_violin() + 
  facet_grid(day_of_tx ~ outcome)

```